---
title: "Creating and Applying an Age-Length Key"
author: "Derek H. Ogle, Northland College"
date: "5-Mar-2015"
output: pdf_document
geometry: margin=0.5in
---

```{r echo=FALSE, results='hide'}
library(knitr)
source("../knitr_setup.R")
```

#Preliminaries
```{r eval=FALSE}
# clears objects in R workspace
rm(list = ls())
```

```{r results='hide', warning=FALSE, message=FALSE}
# load needed packages
library(fishWiDNR)   # for setDBClasses()
library(FSA)         # for lencat(), filterD()
library(dplyr)       # for select(), mutate(), arrange(),%>%
library(magrittr)    # for %<>%
library(lubridate)   # for month()
library(plotrix)     # for plotH(), histStack()
```

# Loading Data and Initial Manipulations
```{r cache=TRUE, results='hide'}
# Load and prepare data ... copied from previous handout
setwd("C:/aaaWork/Web/fishR/Courses/WiDNR_Statewide_2015/Day1_IntroR_FMData")
d <- read.csv("SAWYER_fish_raw_data_012915.csv",stringsAsFactors=FALSE,na.strings=c("-","NA","")) %>%
  setDBClasses(type="RDNR") %>%
  select(County,Waterbody.Name,Survey.Year,Sample.Date,Gear,Fish.Data.Seq.No,Species,
         Length.or.Lower.Length.IN,Gender,Age..observed.annuli.,Edge.Counted.Desc,Age.Structure) %>%
  mutate(mon=month(Sample.Date,label=TRUE)) %>%
  mutate(lcat=lencat(Length.or.Lower.Length.IN,w=0.5)) %>%
  arrange(Species,Length.or.Lower.Length.IN)
```

```{r cache=TRUE, results='hide'}
wae <- filterD(d,Waterbody.Name=="NELSON LAKE",Survey.Year==2014,mon=="May",Species=="WALLEYE",
               Gender!="U",Length.or.Lower.Length.IN>11.5,Length.or.Lower.Length.IN<21)

waeF <- filterD(wae,Gender=="F")
waeM <- filterD(wae,Gender=="M")
```

# Construct an Age-Length Key -- Males
```{r}
waeM.aged <- filterD(waeM,!is.na(Age..observed.annuli.))
waeM.aged$Age..observed.annuli.
```

\newpage
```{r fig.width=5}
( rawM <- xtabs(~lcat+Age..observed.annuli.,data=waeM) )
alkM1 <- prop.table(rawM,margin=1)
print(alkM1,digits=2,zero.print="-")                         # for display only
```

\newpage
```{r fig.width=5}
alkPlot(alkM1,pal="gray",xlab="Total Length (in)")
```

# Apply an Age-Length Key -- Males
```{r}
waeM.unaged <- filter(waeM,is.na(Age..observed.annuli.))
waeM.unaged <- alkIndivAge(alkM1,Age..observed.annuli.~Length.or.Lower.Length.IN,data=waeM.unaged)
waeM.fnl <- rbind(waeM.aged,waeM.unaged)
waeM.fnl$Age..observed.annuli.
```

# Construct and Apply an Age-Length Key -- Females
\vspace{-6pt}
Code is in the script.  However, it is mostly a copy-and-paste of the code from above with the 'M's changed to 'F's
```{r echo=FALSE, results='hide', fig.show='hide', warning=FALSE, cache=TRUE}
waeF.aged <- filter(waeF,!is.na(Age..observed.annuli.))

rawF <- xtabs(~lcat+Age..observed.annuli.,data=waeF)
alkF1 <- prop.table(rawF,margin=1)
print(alkF1,digits=2,zero.print="-")                         # for display only
alkPlot(alkF1,pal="gray",xlab="Total Length (in)")

waeF.unaged <- filter(waeF,is.na(Age..observed.annuli.))
waeF.unaged <- alkIndivAge(alkF1,Age..observed.annuli.~Length.or.Lower.Length.IN,data=waeF.unaged)
waeF.fnl <- rbind(waeF.aged,waeF.unaged)
headtail(waeF.fnl)
```

# Application Assignment
\vspace{-6pt}
Create a script that performs the following tasks:

1. Load your FM data into R.  Filter your data to a waterbody, species, year, and sampling date such that some sampled fish were aged and some were not.  Perhaps, also filter by sex.  [Alternatively, use the Sawyer County database and select Walleye in 2014 or 2013 from Lake Chippewa]
1. Construct an **appropriate** age-length key (ALK) from the aged fish.  Visualize the ALK in both tabular and graphical form.
1. Apply the ALK to the unaged fish.
1. Combine the fish aged from a structure and those with ages estimated from the ALK to form a combined data.frame.
1. Confirm that all fish in the final data.frame have assigned ages.
\vspace{12pt}
1. (*Time Permitting*) Repeat the above for the other sex or another species.

**Save your script!**

```{r echo=FALSE, results='hide'}
# # remove some objects because this will be sourced for 03_LinearModelIntro
rm(alkF1,alkM1,d,rawF,rawM,wae,waeF,waeF.aged,waeF.unaged,waeM,waeM.aged,waeM.unaged)
```
