---
title: "Simple Summaries of Age"
author: "Derek H. Ogle, Northland College"
date: "5-Mar-2015"
output: pdf_document
geometry: margin=0.5in
---

```{r echo=FALSE, results='hide'}
library(knitr)
source("../knitr_setup.R")
```

#Preliminaries
```{r results='hide', warning=FALSE, message=FALSE}
library(fishWiDNR)   # for setDBClasses()
library(FSA)         # for expandCounts()
library(dplyr)       # for %>%
```

#Loading Data and Initial Examination
```{r, cache=TRUE}
setwd("C:/aaaWork/Web/fishR/Courses/WiDNR_Statewide_2015/Day2_AgeGrowth")
d <- read.csv("length_weight_age_raw_data.csv",stringsAsFactors=FALSE,na.strings=c("-","NA","")) %>%
  setDBClasses(type="RDNR") %>%
  select(County,Waterbody.Name,Survey.Year,Gear.Type,Species,Gender,Length.IN,Age..observed.annuli.,Age.Structure,Edge.Counted.Desc) %>%
  rename(waterbody=Waterbody.Name,year=Survey.Year,sex=Gender,age=Age..observed.annuli.,strux=Age.Structure,edge=Edge.Counted.Desc)
head(d)
```

```{r}
dgl <- filter(d,waterbody=="GRINDSTONE LAKE",year==2006,Species=="WALLEYE") %>% droplevels()
head(dgl)
```

```{r}
hist(~age,data=dgl,breaks=3:20,ylim=c(0,30),xlab="Age (yrs)")
hist(age~sex,data=dgl,breaks=3:20,xlab="Age (yrs)",ncol=1,nrow=3)
```

```{r}
plot(Length.IN~age,data=dgl,pch=16,ylab="Total Length (in.)",xlab="Age (yrs)")
pchs <- c(1,19,3)
plot(Length.IN~age,data=dgl,pch=pchs[sex],ylab="Total Length (in.)",xlab="Age (yrs)")
```

```{r}
dgl %<>% group_by(sex,age)
sum <- summarize(dgl,n=n(),mnLen=mean(Length.IN,na.rm=TRUE),sdLen=sd(Length.IN,na.rm=TRUE),
                 minLen=min(Length.IN,na.rm=TRUE),maxLen=max(Length.IN,na.rm=TRUE))
data.frame(sum)
@
  

\newpage

# Application Assignment

Create a script that performs the following tasks:

1. Load your FM data into R.
1. Set the classes on your data.frame.
1. How many rows are in your data.frame?
1. How many total fish are in your data.frame?
1. Expand the counts to individual fish (assign to a new data.frame).
1. How many variables (columns) and individual fish (rows) are in your new data.frame?
1. Display all expanded length measurements.
1. Show all "values" for two other variables of your choice.

**Save your script!**
