---
title: "Summarizing Age Data"
author: "Derek H. Ogle, Northland College"
date: "5-Mar-2015"
output: pdf_document
geometry: margin=0.5in
---

```{r echo=FALSE, results='hide'}
library(knitr)
source("../knitr_setup.R")
```

\vspace{-24pt}
#Preliminaries
\vspace{-12pt}
```{r results='hide', fig.show='hide', message=FALSE, warning=FALSE}
source("02_AgeLengthKey.R")
```

\vspace{-18pt}
```{r}
ls()
```

\vspace{-18pt}
```{r}
hist(~Age..observed.annuli.,data=waeM.fnl,xlab="Age (yrs)",ylab="Male Walleye Captured")

histStack(Length.or.Lower.Length.IN~Age..observed.annuli.,data=waeM.fnl,xlab="Total Length (in.)",
           col="gray.colors",right=FALSE,legend.pos="topright")

waeM.sumlen <- waeM.fnl %>%
  group_by(Age..observed.annuli.) %>%
  summarize(n=n(),mean=mean(Length.or.Lower.Length.IN),sd=sd(Length.or.Lower.Length.IN),
            min=min(Length.or.Lower.Length.IN),max=max(Length.or.Lower.Length.IN))
waeM.sumlen

plotH(n~Age..observed.annuli.,data=waeM.sumlen,xlab="Age (yrs)",ylab="Male Walleye Captured",
       xlim=c(1.5,9.5),width=0.9)
```

\newpage
```{r}
plot(Length.or.Lower.Length.IN~Age..observed.annuli.,data=waeM.fnl,pch=16,col=rgb(0,0,0,1/10),
      xlab="Age",ylab="Total Length (mm)",ylim=c(10,21))
lines(mean~Age..observed.annuli.,data=waeM.sumlen,lwd=2,lty=2)
```

# Construct and Aply an Age-Length Key -- Females
Copy the code from above and convert the 'M's to 'F's


# Application Assignment

Create a script that performs the following tasks:

1. Continue or `source()` your previous script.
1. Summarize the age distribution from the fish in your sample.
1. Show the mean length-at-age for all fish in your sample in both tabular and graphical forms.
1. (*Time Permitting*) Show the length frequency for all fish in your sample.
1. (*Time Permitting*) Repeat the above for your second sex or species.

**Save your script!**
