---
title: "Summarizing Age Data"
author: "Derek H. Ogle, Northland College"
date: "5-Mar-2015"
output: pdf_document
geometry: margin=0.5in
---

```{r echo=FALSE, results='hide'}
library(knitr)
source("../knitr_setup.R")
```

\vspace{-24pt}
#Preliminaries
```{r eval=FALSE}
# clears objects in R workspace
rm(list = ls())
```

```{r results='hide', fig.show='hide', message=FALSE, warning=FALSE}
# sourceing the script that constructed and applied the ALK
# may need to adjust if you named your script differently
source("02_AgeLengthKey.R")
```

```{r echo=FALSE, warning=FALSE}
# ############################################################
# # This is used only for preparation of the handout
rm(hook1)
# ############################################################
```

```{r}
# to demonstrate what is in the workspace after the sourceing
ls()
```

```{r}
hist(~Age..observed.annuli.,data=waeM.fnl,xlab="Age (yrs)",ylab="Male Walleye Captured")
```

\newpage

```{r}
histStack(Length.or.Lower.Length.IN~Age..observed.annuli.,data=waeM.fnl,xlab="Total Length (in.)",
           right=FALSE,legend.pos="topright")

waeM.sumlen <- waeM.fnl %>%
  group_by(Age..observed.annuli.) %>%
  summarize(n=n(),mean=mean(Length.or.Lower.Length.IN),sd=sd(Length.or.Lower.Length.IN),
            min=min(Length.or.Lower.Length.IN),max=max(Length.or.Lower.Length.IN))
waeM.sumlen
```

\newpage
```{r}
plotH(n~Age..observed.annuli.,data=waeM.sumlen,xlab="Age (yrs)",ylab="Male Walleye Captured",
       xlim=c(1.5,9.5),width=0.9)

plot(Length.or.Lower.Length.IN~Age..observed.annuli.,data=waeM.fnl,pch=16,col=rgb(0,0,0,1/10),
      xlab="Age",ylab="Total Length (mm)",ylim=c(10,21))
lines(mean~Age..observed.annuli.,data=waeM.sumlen,lwd=2,lty=2)
```

\newpage
```{r}
plot(Length.or.Lower.Length.IN~Age..observed.annuli.,data=waeM.fnl,pch=16,col=rgb(0,0,0,1/10),
      xlab="Age",ylab="Total Length (mm)",ylim=c(10,21))
lines(mean~Age..observed.annuli.,data=waeM.sumlen,lwd=2,lty=2)

wae.JM <- data.frame(age=1:17,
                      state=c(6.5,9.8,12.0,14.1,16.1,17.8,19.3,20.7,21.8,
                              22.9,23.8,24.5,25.1,25.9,25.5,25.8,25.2),
                      NOR=c(6.4,9.5,11.7,13.8,15.8,17.5,19.1,20.5,21.6,
                            22.7,23.7,24.4,25.2,25.8,25.6,25.6,NA))
lines(state~age,data=wae.JM,lwd=2,lty=2,col="red")
lines(NOR~age,data=wae.JM,lwd=2,lty=2,col="blue")
```

# Construct and Apply an Age-Length Key -- Females
Copy the code from above and convert the 'M's to 'F's


# Application Assignment

Create a script that performs the following tasks:

1. Continue or `source()` your script from the previous handout.
1. Summarize the age distribution from all of the fish in your sample.
1. Show the mean length-at-age for all fish in your sample in both tabular and graphical forms.
\vspace{12pt}
1. (*Time Permitting*) Show the length frequency for all fish in your sample.
1. (*Time Permitting*) Repeat the above for your second sex or species.

**Save your script!**

```{r results='hide', echo=FALSE, fig.show='hide', message=FALSE}
# ############################################################
# ############################################################
# # Create a formatted table as a picture that allows results
# # to be cirectly input into presentations or reports
# #    largely from Gretchen Hansen
# ############################################################

# # Load some extra packages
library(ggplot2)
library(gridExtra)
library(magrittr)

# # Change the name of "age" variable in waeM.sumlen
# # and round the mean and sd
waeM.sumlen %<>% rename(Age=Age..observed.annuli.) %>%
  mutate(mean=round(mean,1),sd=round(sd,2))

# # Create a title for the table and the outputted file
ttl <- "Mean Length-at-Age for Nelson Lake Walleye, Spring 2013"

# # This creates the table (note the use of show.rownames=TRUE)
p <- ggplot(waeM.sumlen,aes(x=1,y=1)) + 
  annotation_custom(tableGrob(waeM.sumlen,show.rownames=FALSE)) +
  ggtitle(ttl) +
  theme_bw() +
  theme(panel.border=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        plot.title=element_text(size=rel(1.2),face="bold",vjust=2),
        axis.ticks=element_blank(), 
        axis.text=element_blank(),
        axis.title=element_blank())
p

ggsave(paste0(ttl,".png"),p,width=6,height=3.5)
# ############################################################
# ############################################################
```
