---
title: "(Very) Quick Introduction to Linear Models in R"
author: "Derek H. Ogle, Northland College"
date: "5-Mar-2015"
output: pdf_document
geometry: margin=0.5in
---

\vspace{-24pt}

# Preliminaries
\vspace{-12pt}
```{r echo=FALSE, results='hide'}
library(knitr)
source("../knitr_setup.R")
```

```{r eval=FALSE}
# clears objects in R workspace
rm(list = ls())
```

```{r results='hide', warning=FALSE, message=FALSE}
# load needed packages
library(FSA)         # for headtail(), filterD(), fitPlot()
library(dplyr)       # for mutate(),%>%
library(multcomp)    # for glht(), mcp()
```

```{r}
ruf <- read.csv("RuffeSLRH.csv") %>%
  mutate(logW=log10(wt),logL=log10(tl),fYear=factor(year))
headtail(ruf)
```

# Simple Linear Regession
\vspace{-12pt}
```{r}
ruf90 <- filterD(ruf,year==1990)
fit1 <- lm(logW~logL,data=ruf90)
coef(fit1)
confint(fit1)
anova(fit1)
```

```{r}
( tmp <- range(ruf90$logL) )
xs <- seq(tmp[1],tmp[2],length.out=99)
xs[1:10]
ys <- predict(fit1,data.frame(logL=xs))
ys[1:10]
```

```{r}
plot(logW~logL,data=ruf90,pch=19,col=rgb(0,0,0,1/6),ylab="log Weight (g)",xlab="log Total Length (mm)")
lines(ys~xs,lwd=2)

plot(wt~tl,data=ruf90,pch=19,col=rgb(0,0,0,1/6),ylab="Weight (g)",xlab="Total Length (mm)")
btxs <- 10^xs
btys <- 10^ys
lines(btys~btxs,lwd=2)
```

\newpage

# Dummy Variable Regession (aka ANCOVA)
```{r}
ruf9000 <- filterD(ruf,year %in% c(1990,2000))
fit2 <- lm(logW~logL*fYear,data=ruf9000)
anova(fit2)
coef(fit2)
confint(fit2)

tmp <- ruf9000 %>% group_by(fYear) %>% summarize(min=min(tl,na.rm=TRUE),max=max(tl,na.rm=TRUE))
tmp

# base plot
clrs <- c(rgb(0,0,0,1/6),rgb(0,0,1,1/6))
plot(wt~tl,data=ruf9000,pch=19,col=clrs[ruf9000$fYear],ylab="Weight (g)",xlab="Total Length (mm)")

# plot line for 1990
tmpx <- seq(tmp$min[1],tmp$max[1],length.out=99)
tmpy <- 10^(predict(fit2,data.frame(logL=log10(tmpx),fYear=factor(1990))))
lines(tmpy~tmpx,lwd=2)

# plot line for 2000
tmpx <- seq(tmp$min[2],tmp$max[2],length.out=99)
tmpy <- 10^(predict(fit2,data.frame(logL=log10(tmpx),fYear=factor(2000))))
lines(tmpy~tmpx,col="blue",lwd=2)

# add a legend
legend("topleft",c("1990","2000"),pch=19,col=c("black","blue"),bty="n")
```

\vspace{36pt}
# 1-way ANOVA
```{r}
ruf2 <- filterD(ruf,year %in% c(1990,1995,2000,2006))
fit3 <- lm(tl~fYear,data=ruf2)
anova(fit3)

mc1 <- glht(fit3,mcp(fYear="Tukey"))
summary(mc1)
cld(mc1)
fitPlot(fit3,ylab="Total Length (mm)",xlab="Capture Year")
```


\vspace{36pt}
# Application Assignment

Create a script that performs the following tasks:

1. Load the `BLGLW.CSV` data into an R data.frame.
1. Determine if there is a significant relationship between the weight and length of Bluegill in these data.
1. Determine if there is a significant difference between sexes in the relationship between the weight and length of Bluegill.
1. Determine if there is a significant difference among the three lakes in the mean weight of Bluegill.

**Save your script!**
