---
title: "Add and Rename Variables"
author: "Derek H. Ogle, Northland College"
date: "4-Mar-2015"
output: pdf_document
geometry: margin=0.6in
---

```{r echo=FALSE, results='hide'}
library(knitr)
source("../knitr_setup.R")
```

#Preliminaries
```{r eval=FALSE}
# clears objects in R workspace
rm(list = ls())
```

```{r results='hide', warning=FALSE, message=FALSE}
# load needed packages
library(fishWiDNR)   # for setDBClasses(), changeDBNames()
library(dplyr)       # for filter(), select(), mutate(), rename()
library(lubridate)   # for month()
library(FSA)         # for expandCounts(), capFirst(), filterD()
```

```{r, cache=TRUE, message=FALSE}
# load FM data and expand lengths ... mostly copied code from first and second handouts
setwd("C:/aaaWork/Web/fishR/Courses/WiDNR_Statewide_2015/Day1_IntroR_FMData")
d <- read.csv("SAWYER_fish_raw_data_012915.csv",stringsAsFactors=FALSE,na.strings=c("-","NA",""))
d <- setDBClasses(d,type="RDNR")
d <- expandCounts(d,~Number.of.Fish,~Length.or.Lower.Length.IN+Length.Upper.IN,new.name="Len")
```

```{r, cache=TRUE, message=FALSE}
d1 <- filter(d,Species=="LAKE STURGEON",Waterbody.Name=="BARKER LAKE",!is.na(Weight.Pounds))
d1 <- select(d1,Species,Survey.Year,Survey.Begin.Date,Len,Weight.Pounds)
headtail(d1,n=2)
```

# Variable Additions
## Simple Mutations

```{r}
tmp <- mutate(d1,loglen=log(Len),logwt=log(Weight.Pounds))
headtail(tmp,n=2)
```

## Simple Special Purpose Mutations

```{r}
tmp <- mutate(d1,mon1=month(Survey.Begin.Date),
                  mon2=month(Survey.Begin.Date,label=TRUE))
headtail(tmp,n=2)
```

```{r}
tmp <- mutate(d1,Species1=capFirst(Species),
                  Species2=capFirst(Species,which="first"))
headtail(tmp,n=2)
```

## Length Category Mutations

```{r}
tmp <- mutate(d1,lcat2=lencat(Len,w=2),
                  lcat2a=lencat(Len,w=2,as.fact=TRUE),
                  lcatA=lencat(Len,breaks=c(46,54,56,58,70)),
                  lcatB=lencat(Len,breaks=c(small=0,medium=50,large=60,very_large=70),use.names=TRUE) )
headtail(tmp)
xtabs(~lcat2,data=tmp)
xtabs(~lcat2a,data=tmp)
xtabs(~lcatA,data=tmp)
xtabs(~lcatB,data=tmp)
```

\newpage

# Rename Variables

```{r}
tmp <- rename(d1,year=Survey.Year,wt=Weight.Pounds)
headtail(tmp)
```

```{r}
tmp <- changeDBNames(d1)
names(tmp)
```

```{r}
tmp <- changeDBNames(tmp,from="R",to="RDNR")
names(tmp)
```

```{r}
tmp <- changeDBNames(tmp,from="RDNR",to="DNR")
names(tmp)
tmp$"Weight Pounds"
write.csv(tmp,"LKS_Barker.csv",row.names=FALSE)
```


\newpage

# Application Assignment

Create a script that performs the following tasks:

1. Load and prepare (set classes, expand counts, examine structure) your FM data in R (**HINT:** *use all or some of your scripts from previous application assignments*).
1. Rename two or more variables to names that better fit your usage (or change all names according to the definitions in `changeDBNames()`).
1. Create a new variable that has the species names with only the first letters capitalized.
1. Create a new variable that has the water body names with only the first letters capitalized.
1. Create a new variable that is the length in mm computed from the length in inches (even though this is already in the FM database).
1. Isolate a game species from a waterbody (and possibly a gear) of interest to you.
1. Create a new variable that contains evenly-spaced length categories that are appropriate for your species.  Construct a frequency table of that variable.
1. Create a new variable that contains length categories that could be defined as "not of interest", "marginally interesting", "preferred", and "very interesting" to anglers for your species.  Construct a frequency table of that variable.
\vspace{12pt}
1. (*Time Permitting*) Create a new variable that contains the Gabelhouse length categories ("stock", "quality", etc.) for your species (**HINT**: *use, for example,* `psdVal("Largemouth Bass",units="in")` *to find Gablehouse lengths for a particular species*).

**Save your script!**
