---
title: "Size Structure I"
author: "Derek H. Ogle, Northland College"
date: "4-Mar-2015"
output: pdf_document
geometry: margin=0.6in
---

```{r echo=FALSE, results='hide'}
fn <- "04_SizeStructure_I.Rmd"
library(knitr)
source("../knitr_setup.R")
```

#Preliminaries
```{r eval=FALSE}
# clears objects in R workspace
rm(list = ls())
```

```{r results='hide', warning=FALSE, message=FALSE}
# load needed packages
library(fishWiDNR)   # for setDBClasses()
library(dplyr)       # for filter(), select(), mutate(), group_by(), summarize()
#  options(dplyr.print_max=1e9)
library(FSA)         # for expandCounts(), filterD(), Summarize(), hist(), 
library(lubridate)   # for month()
```

```{r, cache=TRUE, message=FALSE}
# load FM data, expand lengths, select pertintent variables ... mostly copied code from previous
setwd("C:/aaaWork/Web/fishR/Courses/WiDNR_Statewide_2015/Day1_IntroR_FMData")
d <- read.csv("SAWYER_fish_raw_data_012915.csv",stringsAsFactors=FALSE,na.strings=c("-","NA",""))
d <- setDBClasses(d,type="RDNR")
d <- expandCounts(d,~Number.of.Fish,~Length.or.Lower.Length.IN+Length.Upper.IN,new.name="Len")
d <- mutate(d,Mon=month(Survey.Begin.Date,label=TRUE))
d <- select(d,Species,Waterbody.Name,Survey.Year,Gear,Survey.Begin.Date,Mon,Len)
```

```{r, cache=TRUE, message=FALSE}
Spr <- filterD(d,Survey.Year==2013,Mon %in% c("Apr","May","Jun"))
BGSpr <- filterD(Spr,Species=="BLUEGILL")
BGSprLC <- filterD(BGSpr,Waterbody.Name=="LAKE CHETAC",Gear=="BOOM SHOCKER")
```
So ...

* `Spr` has all species sampled from all water bodies in the Spring of 2013.
* `BGSpr` has only Bluegill sampled from all water bodies in the Spring of 2013.
* `BGSprLC` has only Bluegill sampled with boom shockers from Lake Chetac in the Spring of 2013.

... and they all look similar to this ...

```{r echo=FALSE}
headtail(BGSprLC)
```

\vspace{18pt}

#Very Simple Summaries

```{r}
Summarize(~Len,data=BGSprLC,digits=2)
```

\newpage

#Length Frequency Histograms

```{r hist_basic}
hist(~Len,data=BGSprLC)
```

```{r hist_options}
hist(~Len,data=BGSprLC,xlab="Total Length (In.)",ylab="Number of Bluegill",
     xlim=c(3,9),ylim=c(0,80),col="salmon")
```

\newpage

```{r hist_breaks}
hist(~Len,data=BGSprLC,xlab="Total Length (In.)",ylab="Number of Bluegill",
     xlim=c(3,9),ylim=c(0,40),breaks=seq(3,9,0.2),col="#FA8072")
```

#Multiple Summaries at Once

```{r}
BGSpr <- group_by(BGSpr,Waterbody.Name)
summarize(BGSpr,n=n(),meanLen=mean(Len))                    # see use of na.rm=TRUE below
```

\newpage

```{r}
summarize(BGSpr,n=n(),valid_n=sum(!is.na(Len)),
           meanLen=mean(Len,na.rm=TRUE),sdLen=sd(Len,na.rm=TRUE),
           minLen=min(Len,na.rm=TRUE),maxLen=max(Len,na.rm=TRUE)  )
```

```{r}
BGSpr <- filterD(BGSpr,Len>=3)
summarize(BGSpr,n=n(),valid_n=sum(!is.na(Len)),
           meanLen=round(mean(Len,na.rm=TRUE),2),sdLen=round(sd(Len,na.rm=TRUE),2),
           minLen=min(Len,na.rm=TRUE),maxLen=max(Len,na.rm=TRUE),
           PSDQ=perc(Len,6,digits=0),PSD7=perc(Len,7,digits=0),PSDP=perc(Len,8,digits=0)  )
```

```{r}
Spr <- group_by(Spr,Waterbody.Name,Species)
tmp <- summarize(Spr,n=n(),valid_n=sum(!is.na(Len)),
                 meanLen=round(mean(Len,na.rm=TRUE),2),sdLen=round(sd(Len,na.rm=TRUE),2) )
tmp                                                           # only partial results shown
write.csv(tmp,"LenSum_Sawyer_Spr13.csv",row.names=FALSE)
```

\newpage

# Application Assignment

Create a script that performs the following tasks:

1. Load and prepare your FM data in R (**HINT:** *use all or some of your scripts from previous application assignments*).
1. Reduce your data.frame to one year and several (4 or more) species of interest.  Call this the *original data.frame.*
1. Reduce the *original data.frame* to one water body and one species of interest.
    * Compute summary stastistics for the length variable.
    * Construct a length frequency histogram.
    * Does your description of the length frequency change dramatically with different bin widths?
1. Reduce the *original data.frame* to only one species (this will include several water bodies).
    * Efficiently construct summary statistics for the length variable for each water body.  Include PSD values that are of interest to you  (**HINT**: *use, for example,* `psdVal("Largemouth Bass",units="in")` *to find Gabelhouse lengths for a particular species*).
\vspace{12pt}
1. (*Time Permitting*) Re-create the summary statistics for one species in each water body but include calculations of the median and first and third quartiles (**HINT**: *use, for example,* `quantile(x,0.50,na.rm=TRUE)` *to compute the median (i.e., 50% quantile) of the data in* `x`.).
1. (*Time Permitting*) Compute summary statistics of the length variable for each water body AND each of the several species of interest to you.  Save the summary statistics to an object and write the results to a CSV file.

**Save your script!**

```{r results='hide', echo=FALSE, fig.show='hide', message=FALSE}
# ############################################################
# ############################################################
# # Create a formatted table as a picture that allows results
# # to be cirectly input into presentations or reports
# #    largely from Gretchen Hansen
# ############################################################

# # Load some extra packages
library(ggplot2)
library(gridExtra)

# # Get a summarized data.frame (2nd to last example from above)
tmp <- summarize(BGSpr,n=n(),valid_n=sum(!is.na(Len)),
                 meanLen=round(mean(Len,na.rm=TRUE),2),sdLen=round(sd(Len,na.rm=TRUE),2),
                 minLen=min(Len,na.rm=TRUE),maxLen=max(Len,na.rm=TRUE),
                 PSDQ=perc(Len,6,digits=0),PSD7=perc(Len,7,digits=0),PSDP=perc(Len,8,digits=0)  )

# # Create a title for the table and the outputted file
ttl <- "Summary Statistics Sawyer Co. Bluegill Spring 2013"

# # This creates the table
p <- ggplot(tmp,aes(x=1,y=1)) + 
  annotation_custom(tableGrob(tmp,show.rownames=FALSE)) +
  ggtitle(ttl) +
  theme_bw() +
  theme(panel.border=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        plot.title=element_text(size=rel(1.2),face="bold",vjust=2),
        axis.ticks=element_blank(), 
        axis.text=element_blank(),
        axis.title=element_blank())
p

ggsave(paste0(ttl,".png"),p,width=9,height=4)
# ############################################################
# ############################################################
```

