---
title: "Size Structure II"
author: "Derek H. Ogle, Northland College"
date: "4-Mar-2015"
output: pdf_document
geometry: margin=0.6in
---

```{r echo=FALSE, results='hide'}
library(knitr)
source("../knitr_setup.R")
```

#Preliminaries
```{r results='hide', warning=FALSE, message=FALSE}
library(fishWiDNR)   # for setDBClasses()
library(FSA)         # for Summarize(), hist(), expandCounts()
library(magrittr)    # for %<>%
library(dplyr)       # for %>%, filter(), select(), mutate(), group_by(), summarize()
library(lubridate)   # for month()
```

```{r, cache=TRUE, message=FALSE}
setwd("C:/aaaWork/Web/fishR/Courses/WiDNR_Statewide_2015/Day1_IntroR_FMData")
d <- read.csv("SAWYER_fish_raw_data_012915.csv",stringsAsFactors=FALSE,na.strings=c("-","NA","")) %>%
      setDBClasses(type="RDNR") %>%
      expandCounts(~Number.of.Fish,~Length.or.Lower.Length.IN+Length.Upper.IN,new.name="Len") %>%
      mutate(Mon=month(Survey.Begin.Date,label=TRUE),Species1=capFirst(Species)) %>%
      select(Species,Species1,Waterbody.Name,Survey.Year,Mon,Gear,Len)
```
```{r, cache=TRUE, message=FALSE}
Spr <- filterD(d,Survey.Year==2013,Mon %in% c("Apr","May","Jun"))
BGSpr <- filterD(Spr,Species=="BLUEGILL")
BGSprLC <- filterD(BGSpr,Waterbody.Name=="LAKE CHETAC",Gear=="BOOM SHOCKER")
SprLC <- filterD(Spr,Waterbody.Name=="LAKE CHETAC")
```
So ...

* `Spr` has all species sampled from all water bodies in the Spring of 2013.
* `BGSpr` has only Bluegill sampled from all water bodies in the Spring of 2013.
* `BGSprLC` has only Bluegill sampled with boom shockers from Lake Chetac in the Spring of 2013.
* `SprLC` has all species sampled from Lake Chetac (in all gears) in the Spring of 2013.

... and they all look similar to this ...

```{r echo=FALSE}
head(BGSprLC)
```

\vspace{18pt}

\newpage

#PSDs Single Waterbody and Species
```{r}
brks <- psdVal("Bluegill",units="in",addLens=7)
BGSprLC %<>% mutate(lcat=lencat(Len,breaks=brks),
                     lcat1=lencat(Len,breaks=brks,use.names=TRUE),
                     lcat2=lencat(Len,breaks=brks,use.names=TRUE,drop.levels=TRUE))
head(BGSprLC)
```

```{r}
xtabs(~lcat,data=BGSprLC)
xtabs(~lcat1,data=BGSprLC)
( freq <- xtabs(~lcat2,data=BGSprLC) )
```

```{r}
( rcum <- rcumsum(freq) )
```

```{r}
rcum["stock"]                                                 # demo number of stock fish
rcum/rcum["stock"]*100
```

\newpage

#PSDs for Multiple Waterbodies and Single Species 

```{r}
BGSpr %<>% mutate(lcat2=lencat(Len,breaks=brks,use.names=TRUE,drop.levels=TRUE))
( freq <- xtabs(~Waterbody.Name+lcat2,data=BGSpr) )
apply(freq,MARGIN=1,FUN=rcumsum)                              # apply result has wrong orientation
( rcum <- t(apply(freq,MARGIN=1,FUN=rcumsum)) )
rcum <- rcum[,-1]                                             # remove "substock" column
rcum/rcum[,"stock"]*100
```


#PSDs for Multiple Species in a Single Waterbody

```{r}
SprLC %<>% mutate(lcat2=psdAdd(Len,Species1,units="in"))
head(SprLC)

( freq <- xtabs(~Species+lcat2,data=SprLC) )
( rcum <- t(apply(freq,MARGIN=1,FUN=rcumsum)) )
rcum <- rcum[,-1] 
rcum/rcum[,"stock"]*100
```


# Application Assignment

Create a script that performs the following tasks:

1. Load and prepare your FM data in R (**HINT:** *use all or some of your scripts from previous application assignments*).
1. Compute the PSDs for one species in one waterbody in one year.
1. Compute the PSDs for one species in all waterbodies in one year in your FM data.
1. Compute the PSDs for all species in one waterbody in one year in your FM data.
1. (*Time Permitting*) Repeat any of the above for another year.

**Save your script!**
