---
title: "Size Structure II"
author: "Derek H. Ogle, Northland College"
date: "4-Mar-2015"
output: pdf_document
geometry: margin=0.6in
---

```{r echo=FALSE, results='hide'}
library(knitr)
source("../knitr_setup.R")
```

# Preliminaries
```{r eval=FALSE}
# clears objects in R workspace
rm(list = ls())
```

```{r results='hide', warning=FALSE, message=FALSE}
# load needed packages
library(fishWiDNR)   # for setDBClasses()
library(FSA)         # for expandCounts(), Summarize(), hist()
library(magrittr)    # for %<>%
library(dplyr)       # for %>%, filter(), select(), mutate(), group_by(), summarize()
#  options(dplyr.print_max=1e9)
library(lubridate)   # for month()
```

```{r, cache=TRUE, message=FALSE}
# load FM data, expand lengths, add variables, select pertintent variables
setwd("C:/aaaWork/Web/fishR/Courses/WiDNR_Statewide_2015/Day1_IntroR_FMData")
d <- read.csv("SAWYER_fish_raw_data_012915.csv",stringsAsFactors=FALSE,na.strings=c("-","NA","")) %>%
      setDBClasses(type="RDNR") %>%
      expandCounts(~Number.of.Fish,~Length.or.Lower.Length.IN+Length.Upper.IN,new.name="Len") %>%
      mutate(Mon=month(Survey.Begin.Date,label=TRUE),Species1=capFirst(Species)) %>%
      select(Species,Species1,Waterbody.Name,Survey.Year,Mon,Gear,Len)
```
```{r, cache=TRUE, message=FALSE}
# create some subsets for use below.
Spr <- filterD(d,Survey.Year==2013,Mon %in% c("Apr","May","Jun"))
BGSpr <- filterD(Spr,Species=="BLUEGILL")
BGSprLC <- filterD(BGSpr,Waterbody.Name=="LAKE CHETAC",Gear=="BOOM SHOCKER")
SprLC <- filterD(Spr,Waterbody.Name=="LAKE CHETAC")
```

Quick questions

* What variables should be in `Spr`, `BGSpr`, `BGSprLC`, and `SprLC`.
* What individuals should be in `Spr`, `BGSpr`, `BGSprLC`, and `SprLC`.

\newpage

#PSDs Single Waterbody and Species
```{r}
( brks <- psdVal("Bluegill",units="in",addLens=7) )
BGSprLC %<>% mutate(lcat=lencat(Len,breaks=brks),
                     lcat1=lencat(Len,breaks=brks,use.names=TRUE),
                     lcat2=lencat(Len,breaks=brks,use.names=TRUE,drop.levels=TRUE))
headtail(BGSprLC)
```

```{r}
xtabs(~lcat,data=BGSprLC)
xtabs(~lcat1,data=BGSprLC)
( freq <- xtabs(~lcat2,data=BGSprLC) )
```

```{r}
( rcum <- rcumsum(freq) )
```

```{r}
rcum["stock"]                                                 # demo number of stock fish
rcum/rcum["stock"]*100
```

\newpage

#PSDs for Multiple Waterbodies and Single Species 

```{r}
BGSpr %<>% mutate(lcat2=lencat(Len,breaks=brks,use.names=TRUE,drop.levels=TRUE))
( freq <- xtabs(~Waterbody.Name+lcat2,data=BGSpr) )
```
```{r, results='hide'}
apply(freq,MARGIN=1,FUN=rcumsum)     # apply result has wrong orientation, only partial results shown
```
```{r, echo=FALSE}
apply(freq,MARGIN=1,FUN=rcumsum)[1:5,1:6]
```

```{r}
( rcum <- t(apply(freq,MARGIN=1,FUN=rcumsum)) )
rcum <- rcum[,-1]                                             # remove "substock" column
rcum/rcum[,"stock"]*100
```


#PSDs for Multiple Species in a Single Waterbody

```{r}
SprLC %<>% mutate(lcat2=psdAdd(Len,Species1,units="in"))
headtail(SprLC)

( freq <- xtabs(~Species+lcat2,data=SprLC) )
( rcum <- t(apply(freq,MARGIN=1,FUN=rcumsum)) )
rcum <- rcum[,-1] 
rcum/rcum[,"stock"]*100
```


# Application Assignment

Create a script that performs the following tasks:

1. Load and prepare your FM data in R (**HINT:** *use all or some of your scripts from previous application assignments*).
1. Compute the PSDs for one species in one waterbody in one year.
1. Compute the PSDs for one species in all waterbodies in one year in your FM data.
1. Compute the PSDs for all species in one waterbody in one year in your FM data.
\vspace{12pt}
1. (*Time Permitting*) Repeat any of the above for another year.

**Save your script!**

```{r results='hide', echo=FALSE, fig.show='hide', message=FALSE}
# ############################################################
# ############################################################
# # Create a formatted table as a picture that allows results
# # to be cirectly input into presentations or reports
# #    largely from Gretchen Hansen
# ############################################################

# # Load some extra packages
library(ggplot2)
library(gridExtra)

# # Save last result from above, must be coerced to a data.frame
tmp <- data.frame(round(rcum/rcum[,"stock"]*100,0))

# # Create a title for the table and the outputted file
ttl <- "PSD Analysis, Lake Chetac, Spring 2013"

# # This creates the table (note the use of show.rownames=TRUE)
p <- ggplot(tmp,aes(x=1,y=1)) + 
  annotation_custom(tableGrob(tmp,show.rownames=TRUE)) +
  ggtitle(ttl) +
  theme_bw() +
  theme(panel.border=element_blank(),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        plot.title=element_text(size=rel(1.2),face="bold",vjust=2),
        axis.ticks=element_blank(), 
        axis.text=element_blank(),
        axis.title=element_blank())
p

ggsave(paste0(ttl,".png"),p,width=6,height=4)
# ############################################################
# ############################################################
```
